---
title: "CUNY MSDS 607 Project 1"
author: "AJ Bentley"
date: "September 20, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(stringr)

```


#### In this project, we were given a text file with chess tournament results where the information has some structure. My job was to create an R Markdown file that generates a .CSV file with the following information for all of the players:
- Player’s Name
- Player’s State
- Total Number of Points
- Player’s Pre-Rating
- Average Pre Chess Rating of Opponents

<br>
Load data into R data frame

```{r load data} 

mydata <- read.table("https://raw.githubusercontent.com/ajbentley/cuny_ms_ds/master/607/tournamentinfo.txt", header = TRUE, skip = 1, sep = "|", fill=TRUE)


```

#### Plan of attack
__Create vectors for each of the requested data (Player's Name, Player's State, Total Number of Points, Player's Pre-Rating, and Average Pre Chess Rating of Opponents), then merge them into a dataframe.__


```{r name vector}

namevector <- mydata[['Player.Name']]

name <- namevector[seq(3, length(namevector), 3)]

dfs <- data.frame(name)
head(dfs)

```

```{r state vector}

statevector <- mydata[['Pair']]

state <- statevector[seq(4, length(statevector), 3)]

dfs$state <- state
head(dfs)

```


```{r player points vector}

pointsvector <- mydata[['Total']]

points <- pointsvector[seq(3, length(statevector), 3)]

dfs$points <- points
head(dfs)

```

```{r player pre rating vector}

playerprerating <- namevector[seq(4, length(namevector), 3)]

plypredf <- data.frame(playerprerating)

plypredf$plypre <- sub('.* R: ', '', plypredf$playerprerating)

plypredf$plypre <- sub('(.{4})(.*)', '\\1', plypredf$plypre)

head(plypredf)

```



```{r player number}

playernumber <- statevector[seq(3, length(statevector), 3)]
```


```{r collect rows up until now}

dfs$playernum <- as.numeric(playernumber)

dfs <- dfs[c("playernum", "name", "state", "points")]

dfs$playerrtg <- plypredf$plypre

head(dfs)

```

Now I've got to add columns with each competitor for each player

```{r competitors}
r1vector <- mydata[['Round']]
rnd1 <- r1vector[seq(3, length(r1vector), 3)]
rnd1 <- str_extract(rnd1, "\\d+")

r2vector <- mydata[['Round.1']]
rnd2 <- r2vector[seq(3, length(r2vector), 3)]
rnd2 <- str_extract(rnd2, "\\d+")

r3vector <- mydata[['Round.2']]
rnd3 <- r3vector[seq(3, length(r3vector), 3)]
rnd3 <- str_extract(rnd3, "\\d+")

r4vector <- mydata[['Round.3']]
rnd4 <- r4vector[seq(3, length(r4vector), 3)]
rnd4 <- str_extract(rnd4, "\\d+")

r5vector <- mydata[['Round.4']]
rnd5 <- r5vector[seq(3, length(r5vector), 3)]
rnd5 <- str_extract(rnd5, "\\d+")

r6vector <- mydata[['Round.5']]
rnd6 <- r6vector[seq(3, length(r6vector), 3)]
rnd6 <- str_extract(rnd6, "\\d+")

r7vector <- mydata[['Round.6']]
rnd7 <- r7vector[seq(3, length(r7vector), 3)]
rnd7 <- str_extract(rnd7, "\\d+")

dfs$rnd1 <- as.numeric(rnd1)
dfs$rnd2 <- as.numeric(rnd2)
dfs$rnd3 <- as.numeric(rnd3)
dfs$rnd4 <- as.numeric(rnd4)
dfs$rnd5 <- as.numeric(rnd5)
dfs$rnd6 <- as.numeric(rnd6)
dfs$rnd7 <- as.numeric(rnd7)

head(dfs)
```

Make relatinal df of just the playernumber and playerrating


```{r make relational db of number and rtg}
df_pr <- data.frame(as.numeric(dfs$playernum), as.numeric(dfs$playerrtg))
head(df_pr)


```

Now merge df_pr with dfs on dfs.playernum matching to each rnd column.

```{r relate dfs}

#merge the two dayparts on rnd1
dfsr1 <- merge(dfs, df_pr, by.x = "rnd1", by.y = "as.numeric.dfs.playernum.", all.x = TRUE)
dfsr1 <- merge(dfsr1, df_pr, by.x = "rnd2", by.y = "as.numeric.dfs.playernum.", all.x = TRUE, suffix = c(".x",".y"))

# have to clean up titles as we go along
colnames(dfsr1)[colnames(dfsr1)=="as.numeric.dfs.playerrtg..y"] <- "rnd1_oppscore"
colnames(dfsr1)[colnames(dfsr1)=="as.numeric.dfs.playerrtg..x"] <- "rnd2_oppscore"

dfsr1 <- merge(dfsr1, df_pr, by.x = "rnd3", by.y = "as.numeric.dfs.playernum.", all.x = TRUE, suffix = c(".x",".y"))
colnames(dfsr1)[colnames(dfsr1)=="as.numeric.dfs.playerrtg."] <- "rnd3_oppscore"

dfsr1 <- merge(dfsr1, df_pr, by.x = "rnd4", by.y = "as.numeric.dfs.playernum.", all.x = TRUE, suffix = c(".x",".y"))
colnames(dfsr1)[colnames(dfsr1)=="as.numeric.dfs.playerrtg."] <- "rnd4_oppscore"

dfsr1 <- merge(dfsr1, df_pr, by.x = "rnd5", by.y = "as.numeric.dfs.playernum.", all.x = TRUE, suffix = c(".x",".y"))
colnames(dfsr1)[colnames(dfsr1)=="as.numeric.dfs.playerrtg."] <- "rnd5_oppscore"

dfsr1 <- merge(dfsr1, df_pr, by.x = "rnd6", by.y = "as.numeric.dfs.playernum.", all.x = TRUE, suffix = c(".x",".y"))
colnames(dfsr1)[colnames(dfsr1)=="as.numeric.dfs.playerrtg."] <- "rnd6_oppscore"

dfsr1 <- merge(dfsr1, df_pr, by.x = "rnd7", by.y = "as.numeric.dfs.playernum.", all.x = TRUE, suffix = c(".x",".y"))
colnames(dfsr1)[colnames(dfsr1)=="as.numeric.dfs.playerrtg."] <- "rnd7_oppscore"

head(dfsr1)


```

Add a column averaging the scores

```{r avg scores}

dfsr1$avg_opponent_score <- rowMeans(dfsr1[c(13:19)], na.rm=TRUE)
head(dfsr1)



```

Order in the data frame!

```{r reorder and reset row names}

dfsr1 <- dfsr1[order(dfsr1$playernum),]
rownames(dfsr1) <- 1:nrow(dfsr1)

head(dfsr1)
```


Time for cleanup!

```{r remove vestigal columns}

#removing columns not needed
dfsr1 <- dfsr1[,-c(1:8)]
dfsr1 <- dfsr1[,-c(5:11)]

```

```{r make avg an int}

# making the average opponent score an integer, as intdicated in the assignment
dfsr1$avg_opponent_score <- as.integer(dfsr1$avg_opponent_score)

```

Finished product

```{r show data}
dfsr1
```

Export to CSV

```{r export}

write.csv(dfsr1, file = "chess_data.csv")
```



