---
title: "CUNY MSDS 607 Lab Week 3"
author: "AJ Bentley"
date: "September 14, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(stringr)

```

#### Q3. The Simpsons
Load data into R data frame

```{r load data} 

raw.data <-"555-1239Moe Szyslak(636) 555-0113Burns, C. Montgomery555-6542Rev. Timothy Lovejoy555 8904Ned Flanders636-555-3226Simpson, Homer5553642Dr. Julius Hibbert"

raw.data
```

<br>

Create name vector 
_(used code from book)_

```{r name charvar}

name <- unlist(str_extract_all(raw.data, "[[:alpha:]., ]{2,}"))

```

<br>

__Translate name list to first_name last_name__

Pseudocode

1. for any item in "name" that contains a ", " after a word and before another word, reverse the order of the words before and after.

2. strip out any titles. be careful of Chuck Burns--that C. is an initial and not a title.


```{r split names on comma}

namefl <- str_replace_all(name, "([[:alpha:]]+)(, )([[:alpha:]].+)", "\\3 \\1")

namefl2 <- str_replace_all(namefl, "[[:alpha:]]{2,}\\.", "")

namefl2
```

<br>

__Construct a logical vector indicating whether a character has a title.__

```{r title?}

str_detect(name, "[[:alpha:]]{2,}\\.")

```

<br>

__Construct a logical vector indicating whether a character has a second name.__

_I am assuming this means a middle name, though the names given exclude Homer and Dr. Hibbert's middle names (or that Rev. Lovejoy and Ned Flanders are both Jr.s)_


```{r middle name}
str_detect(namefl2, "(\\.\\ )\\w*\\ \\w*")

```

<br>

#### Q4. Describe the types of strings that conform to the following regular expressions and construct an example that is matched by the regular expression.

__1. [0-9]+\\$__

    The end of the string is at least 1 number (e.g., 6541 or the last four digits of "asdf.4324")

<br>

__2. \\b[az]{1,4}\\b__

    Any combination of the letters a and z that is less than 5 characters long (e.g., 'a', 'az', 'zaaz', 'aaax')

<br>

__3. .*?\\.txt$__

    Any string that ends in '.txt' (e.g., 'bobsyouruncle.txt')

<br>

__4. \\d{2}/\\d{2}/\\d{4}__

    2 digits followed by a slash followed by 2 digits followed by a slash followed by 4 digits (e.g., 10/23/1973)

<br>

__5. <(.+?)>.+?</\\1>__

    Any number of non-lineterminating characters enclosed in <>, followed by any number of non-lineterminating characters, followed by a slash and the contents of the first set of non-lineterminating characters enclosed in <> (e.g. "<hard>thingsare</hard>")

    
    
    
    