---
title: 'Inference for numerical data'
author: "A. Joshua Bentley"
output:
  html_document:
    css: ./lab.css
    highlight: pygments
    theme: cerulean
  pdf_document: default
---

## North Carolina births

#### In 2004, the state of North Carolina released a large data set containing 
information on births recorded in this state. This data set is useful to 
researchers studying the relation between habits and practices of expectant 
mothers and the birth of their children. We will work with a random sample of 
observations from this data set.

## Exploratory analysis

Load the `nc` data set into our workspace.

```{r load-data, eval=TRUE}
load("more/nc.RData")
```

We have observations on 13 different variables, some categorical and some 
numerical. The meaning of each variable is as follows.

variable         | description                                                               
---------------- | -------------------------------------------------------------------------
`fage`           | father's age in years.
`mage`           | mother's age in years.
`mature`         | maturity status of mother.
`weeks`          | length of pregnancy in weeks.
`premie`         | whether the birth was classified as premature (premie) or full-term.
`visits`         | number of hospital visits during pregnancy.
`marital`        | whether mother is `married` or `not married` at birth.
`gained`         | weight gained by mother during pregnancy in pounds.
`weight`         | weight of the baby at birth in pounds.
`lowbirthweight` | whether baby was classified as low birthweight (`low`) or not (`not low`).
`gender`         | gender of the baby, `female` or `male`.
`habit`          | status of the mother as a `nonsmoker` or a `smoker`.
`whitemom`       | whether mom is `white` or `not white`.

1.  What are the cases in this data set? How many cases are there in our sample?

    The cases are births and there are 1000 cases (ran dim on nc to get to this number).

As a first step in the analysis, we should consider summaries of the data. This
can be done using the `summary` command:

```{r summary, eval=TRUE}
summary(nc)
```

As you review the variable summaries, consider which variables are categorical 
and which are numerical. For numerical variables, are there outliers? If you 
aren't sure or want to take a closer look at the data, make a graph.

Consider the possible relationship between a mother's smoking habit and the 
weight of her baby. Plotting the data is a useful first step because it helps 
us quickly visualize trends, identify strong associations, and develop research
questions.

2.  Make a side-by-side boxplot of `habit` and `weight`. What does the plot 
highlight about the relationship between these two variables?

```{r habit weight boxes}

boxplot(nc$weight~nc$habit)

```

The box plots show how the medians of the two distributions compare, but we can
also compare the means of the distributions using the following function to 
split the `weight` variable into the `habit` groups, then take the mean of each
using the `mean` function.

```{r by-means, eval=TRUE}
by(nc$weight, nc$habit, mean)
```

There is an observed difference, but is this difference statistically 
significant? In order to answer this question we will conduct a hypothesis test
.

## Inference

3.  Check if the conditions necessary for inference are satisfied. Note that 
you will need to obtain sample sizes to check the conditions. You can compute 
the group size using the same `by` command above but replacing `mean` with 
`length`.

```{r by-length, eval=TRUE}
by(nc$weight, nc$habit, length)
```


```{r check distributions}

set.seed(23)
weight_means500 <- rep(NA, 5000)

for(i in 1:5000){
   wsamp <- sample(nc$weight, 500)
   weight_means500[i] <- mean(wsamp)
   }

hist(weight_means500, breaks = 25)

```

    There are more than 30 cases, the weight data resembles a normal curve, and cases are independent so conditions for inference are met.



4.  Write the hypotheses for testing if the average weights of babies born to 
smoking and non-smoking mothers are different.

    H~0~ = The average weights of babies born to smoking and non-smoking mothers are the same.
    
    H~A~ = The average weights of babies born to smoking and non-smoking mothers are different.


Next, we introduce a new function, `inference`, that we will use for conducting
hypothesis tests and constructing confidence intervals. 

```{r inf-weight-habit-ht, eval=TRUE, tidy=FALSE}
inference(y = nc$weight, x = nc$habit, est = "mean", type = "ht", null = 0, 
          alternative = "twosided", method = "theoretical")

```

Let's pause for a moment to go through the arguments of this custom function. 
The first argument is `y`, which is the response variable that we are 
interested in: `nc$weight`. The second argument is the explanatory variable, 
`x`, which is the variable that splits the data into two groups, smokers and 
non-smokers: `nc$habit`. The third argument, `est`, is the parameter we're 
interested in: `"mean"` (other options are `"median"`, or `"proportion"`.) Next
we decide on the `type` of inference we want: a hypothesis test (`"ht"`) or a 
confidence interval (`"ci"`). When performing a hypothesis test, we also need 
to supply the `null` value, which in this case is `0`, since the null 
hypothesis sets the two population means equal to each other. The `alternative`
hypothesis can be `"less"`, `"greater"`, or `"twosided"`. Lastly, the `method` 
of inference can be `"theoretical"` or `"simulation"` based.

5.  Change the `type` argument to `"ci"` to construct and record a confidence 
interval for the difference between the weights of babies born to smoking and 
non-smoking mothers.

```{r inf-weight-habit-ci, eval=TRUE, tidy=FALSE}
inference(y = nc$weight, x = nc$habit, est = "mean", type = "ci", method = "theoretical")

```

   *I don't see any difference between the results of the ci and ht plots*
    


## On your own

-   Calculate a 95% confidence interval for the average length of pregnancies 
(`weeks`) and interpret it in context. Note that since you're doing inference 
on a single population parameter, there is no explanatory variable, so you can 
omit the `x` variable from the function.

```{r inf-weeks-ci, eval=TRUE, tidy=FALSE}
inference(y = nc$weeks, est = "mean", type = "ci", method = "theoretical")

```



-   Calculate a new confidence interval for the same parameter at the 90% 
confidence level. You can change the confidence level by adding a new argument 
to the function: `conflevel = 0.90`.

```{r inf-weeks-ci ci90, eval=TRUE, tidy=FALSE}
inference(y = nc$weeks, est = "mean", type = "ci", method = "theoretical", conflevel = 0.90)

```

-   Conduct a hypothesis test evaluating whether the average weight gained by 
younger mothers is different than the average weight gained by mature mothers.

```{r inf-weight-age-ci, eval=TRUE, tidy=FALSE}
inference(y = nc$weight, x = nc$mature, est = "mean", type = "ht", null = 0, 
          alternative = "twosided", method = "theoretical", 
          order = c("younger mom","mature mom"))
```


-   Now, a non-inference task: Determine the age cutoff for younger and mature 
mothers. Use a method of your choice, and explain how your method works.

    I'll make new dfs filtering by mature factors, then run summaries on each. I know I could just do max on younger and min on older, but I like to see more data.
    
```{r younger mature mom cutoff}

library(dplyr)

ncagemat <- select(nc, mage:mature) 

younger <- filter(ncagemat, mature == "younger mom")
mature <-  filter(ncagemat, mature == "mature mom")

summary(younger)

summary(mature)

```
    
    Younger moms are from age 13 - 34 while mature moms are from 35 - 50. 
    
    While I understand that this is a way to differentiate periods in a woman's life where she is most likely to have a successful, healthy birth, I feel like the labels are both misleading (I believe that maturity is more closely associated with being "of birthing age" than being past it) and presumptive, since it provides an artificial determinant. It would make more sense to create factors based on the data.


-   Pick a pair of numerical and categorical variables and come up with a 
research question evaluating the relationship between these variables. 
Formulate the question in a way that it can be answered using a hypothesis test
and/or a confidence interval. Answer your question using the `inference` 
function, report the statistical results, and also provide an explanation in 
plain language.

__a) Do male children weigh the same as female children?__ 

    H~0~ = The average weights of male children is the same as for female children.
    
    H~A~ = The average weights of male and female babies are different.


```{r inf-weight-gender-ht, eval=TRUE, tidy=FALSE}
inference(y = nc$weight, x = nc$gender, est = "mean", type = "ht", null = 0, 
          alternative = "twosided", method = "theoretical", 
          order = c("male","female"))
```

The male mean is 7.3 while female mean is 6.9, a difference of 0.4. The SD for each gender is about 1.5, which seems pretty high and I'm guessing that is because premies are included so we're going to have to first do the analysis on full term babies and then on premies. 


__b) Do full term male children weigh the same as full term female children?__ 

    H~0~ = The average weights of full term male children is the same as for full term female children.
    
    H~A~ = The average weights of full term male and full term female babies are different.

```{r full term weight by gender}

ncterm <- select(nc, c(premie, weight, gender)) 

fterm <- filter(ncterm, premie == "full term")
premie <- filter(ncterm, premie == "premie")

inference(y = fterm$weight, x = fterm$gender, est = "mean", type = "ht", null = 0, 
          alternative = "twosided", method = "theoretical", 
          order = c("male","female"))
```

Still very surprised and confused by this result. 

Just for giggles I'm going to change the null value to .5 to make sure it's working right.

```{r full term weight by gender .5 test}

inference(y = fterm$weight, x = fterm$gender, est = "mean", type = "ht", null = .5, 
          alternative = "twosided", method = "theoretical", 
          order = c("male","female"))
```

Okay, good. So there's basically no chance that they are the same but it's almost certain that they are within half a pound of each other. This also means that the SD wasn't just a function of premies being mixed with full term babies, but since I'd already said I'd wanted to look at that...

__c) Do premie male children weigh the same as premie female children?__ 

    H~0~ = The average weights of premie male children is the same as for premie female children.
    
    H~A~ = The average weights of premie male and premie female babies are different.

```{r premie weight by gender}

inference(y = premie$weight, x = premie$gender, est = "mean", type = "ht", null = 0, 
          alternative = "twosided", method = "theoretical", 
          order = c("male","female"))
```

The means are much closer, the SDs are larger, and there are no outliers so there is a better chance that there is no difference between males and females in terms of weight in premies.

<div id="license">
This is a product of OpenIntro that is released under a [Creative Commons 
Attribution-ShareAlike 3.0 Unported](http://creativecommons.org/licenses/by-sa/3.0).
This lab was adapted for OpenIntro by Mine &Ccedil;etinkaya-Rundel from a lab 
written by the faculty and TAs of UCLA Statistics.
</div>
